---
title: "Project 1"
author: "Chris Harris / Stephan Kostreski"
date: "March 12, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1 - Warm Up

## Part A

**Generate a sample of 1000 from the normal distribution with a mean of 10 and a standard deviation of 2.25.**

*Hint: Type ?rnorm into the console to see how to use this function to solve this part of the problem. Also note the arguments required:*

* n: number of observations (set to 1000)
* mean: the mean value of the distribution (set to 10)
* standard deviation: the deviation of values from the mean (set to 2.25)

*NOTE: Don't forget to assign the function result to a variable (using the <- operator) so you can refer to it later, like so:*

data_dist <- rnorm(n=10, mean=1, sd=2)


```{r problem_1a}



```

## Part B

**Create a histogram of the data with 10 bins (classes)**

Functions/Arguments needed:

* hist
    + x: data to be tallied (i.e. the variable where you stored the rnorm function call)
    + xlab: label for the x-axis
    + breaks: number of bins (classes) to categorize data
  
*Note: R may not create exactly 10 bins based on its algorithm. To gain precision you may provide exact breaks by using breaks= seq(0,20,by=2). This is what we did:*

*hist(x=data_dist, xlab="Sample Values", breaks=seq(0,20, by=2))*



```{r problem_1b}



```

*At this point, we recommend playing with different parameters to obtain colors, different titles and more. Hint: use ?hist to learn the parameters and/or google*


## Part C

**Create a normal probability plot and boxplot.**

Functions/Arguments needed:

* boxplot - creates a boxplot
    + x: data to be plotted
* qqnorm - creates normal probability plot
    + y: data to be plotted
* qqline - adds a reference line to the qqnorm plot
    + y: data to add reference line
    + col: line color - col=2 creates a red line, col=3 creates a green line
    
    
  
```{r problem_1c}




```

\newpage

# Problem 2

## Part A

Create 1000 samples of size 10.

Functions/Arguments needed:

* rnorm - same parameters as before
* replicate
    + n: number of repititions
    + expr: function to repeat 'n' times
* sample - takes a sample of a particular size from a "population distribution"
    + x: distribution to be sampled from (variable result of rnorm)
    + size: size of sample
    + replace: FALSE (sample without replacement)


```{r problem_2a}



```

## Part B

Calculate the mean for these 1000 and graph the sampling distribution of the sample mean for these samples of size 10.

*There are several ways to accomplish this, but we recommend using a looping construct to take the average of each sample and store each average in a list.*

*Using this method, you should wind up having a list of size 1000. Each element containing the average of that particular sample*

Logical Constructs needed:
  
* For Loop

Functions/Arguments needed:

* c : creates a list with a specified argument range
* mean
    + x: list (function will take the average of this list)
* hist


As this can be tricky for those new to programming, here is an example of how to accomplish this task:

1. We define our distribution population as so:

*distribution <- rnorm(n=100, mean=10, sd=2.25)*

2. We take 10 random samples of size 10 from this distribution

*samples <- replicate(n=10, sample(x=distribution, size=10, replace=FALSE))*

3. We create a new list that will eventually hold our sample averages:

*mean_per_sample <- c(1:10)  - Creates a list of size 10*

4. We then go through each of our individual samples, get the mean of the 10 values in each sample, and store that result in our list.

*for (index in 1:10) { mean_per_sample[index] <- mean(samples[,index]) } *


```{r problem_2b}



```

## Part C

Create a normal probability plot and boxplot of this sampling distribution.

```{r problem_2c}



```

## Part D: Create the 90, 95 and 99 % confidence intervals.

**1. Get the mean, standard deviation, and number of samples taken**


Functions needed:

* mean
* sd
* length


```{r problem_2d_1}



```


**2. Calculate 90% CONFIDENCE INTERVAL:**

We are going to use the standard CI forumla.

Functions needed:

* qnorm - gives us the z-score given a probility
    + p: probability
    
Example:

  qnorm(0.95) = 1.644854
  qnorm(0.975) = 1.959964
  

```{r problem_2d_2}



```

90% CI: []


**3. Calculate 95% CONFIDENCE INTERVAL:**


```{r problem_2d_3}



```

95% CI: []


**4. Calculate 99% CONFIDENCE INTERVAL:**


```{r problem_2d_4}



```

99% CI: []

\newpage

# Problem 3

## Part A

Generate a sample of size 1000 from the exponential with mean 2 

Functions/Arguments needed:

* rexp : randomly generates values from the exponential distribution
    + n: number of values to be generated
    + rate: rate of exponential growth (mean = 1/rate)
    

```{r problem_3a}



```

## Part B

Create a histogram of the data with 10 bins(classes)


```{r problem_3b}



```

## Part C

Create a normal probability plot and boxplot.


```{r problem_3c}



```

## Part D

### 1. Create 1000 samples of size 10.


```{r problem_3d_1}



```

### 2. Calculate the mean for these 1000 and graph the sampling distribution of the sample mean for these samples of size 10


```{r problem_3d_2}



```

### 3. Create the histogram of this data with 10 bins.


```{r problem_3d_3}




```

### 4. Create a normal probability plot and boxplot of this sampling distribution.


```{r problem_3d_4}



```

## Part E

### 1. Create 1000 samples of size 40


```{r problem_3e_1}



```

### 2. Calculate the mean for these 1000 and graph the sampling distribution of the sample mean for these samples of size 40


```{r problem_3e_2}



```

### 3. Create the histogram of this data with 10 bins.


```{r problem_3e_3}




```

### 4. Create a normal probability plot and boxplot of this sampling distribution.


```{r problem_3e_4}



```

### 5. Create the 90, 95 and 99 % confidence intervals.


```{r problem_3e_5}



```

90% Confidence Interval []

95% Confidence Interval []

99% Confidence Interval []

```

